---
title: "Growth Trials"
author: "Lilith Diener"
date: "2025-08-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
options(repos = c(CRAN = "https://cloud.r-project.org"))

```

```{r, warning=FALSE, message=FALSE}

library(gt)
library(tidyr)
library(dplyr)
library(readxl)
library(stringr)
library(dae)
library(ggplot2)
library(lme4)
library(lmerTest)
library(lubridate)
library(emmeans)
library(tidyverse)
library(styler)
library(ggpmisc)
library(car)
library(nlme)
library(tibble)
library(htmltools)

```

Reading in the data and setting the control treatment as baseline level

```{r}

# Read in data
final_data <- read_excel("C:/Users/Win10/Desktop/UCT/Honours Project Gracilaria/Data/Statistical_Analyses_R/Raw_data_complete.xlsx")


# Set the Control treatment as the reference/baseline level
final_data$Treatment <- as.factor(final_data$Treatment)
final_data$Treatment <- relevel(final_data$Treatment, ref = "Control")

```

## *Gracilaria gracilis* 

### Length

Filtering the data and calculating averages

```{r}

# Filter data by species and length
gg_length <- final_data %>%
  filter(Measurement_Type == "Length", Species == "Gracilaria gracilis" )
  
# Calculate mean, SD and SE
gg_avg_length <- gg_length %>%
  group_by(Treatment, Tank_No,Week) %>%
  summarise(Average = mean(Value,na.rm = TRUE), SD = sd(Value), N = n(),
SE = SD / sqrt(N), Date = max(Date),.groups = "drop") %>% 
  ungroup() 

# Average treatment length per week
gg_avg_length_per_week_by_treatment <- gg_avg_length  %>%
  group_by( Week,Treatment) %>%
  summarise(
    Average = mean(Average,na.rm = TRUE), N = n(),
    SD = mean(SD,na.rm = TRUE), SE = mean(SE,na.rm = TRUE),
    Date = max(Date), .groups = "drop")

```

Testing ANOVA assumptions, creating models and running post hoc test

```{r}

# Shapiro-Wilk normality test to test normality of gg avg length per treatment per week
shapiro.test(gg_avg_length_per_week_by_treatment$Average) # normality assumption met

# Levene's Test to check for homogeneity of treatments in gg avg length per treatment per week
leveneTest(gg_avg_length_per_week_by_treatment$Average, 
gg_avg_length_per_week_by_treatment$Treatment, center = mean ) # assumption of equal variances met

# Model of average length per treatment per week
mGAL <-  aov(Average ~ Treatment + Week, data = gg_avg_length_per_week_by_treatment)
summary(mGAL)
aov(mGAL)
anova(mGAL)

# Post hoc test
emmeans_mGAL <- emmeans(mGAL, pairwise ~ Treatment)
emmeans_mGAL$emmeans
emmeans_mGAL$contrasts

```

Figures

```{r, echo=FALSE, warning=FALSE}

# Box plot of distribution of length per treatment per week by tank number
ggplot(gg_avg_length, aes(x = factor(Week), y = Average)) +
  geom_boxplot(position = position_dodge(width = 0.6), alpha = 0.5, fill = "white") +
  facet_wrap(~ Treatment, scales = "free_x") +
  labs(
  title = expression("Mean length of " * italic("Gracilaria gracilis") * " by treatment and week"),
    y = "Mean length (cm)",
    x = "Week"
  ) +
  theme_bw() +  # black-and-white theme
  theme(
    panel.grid = element_blank(),           # remove grid lines
    axis.line = element_line(color = "black") # black x and y axes
  )

# Line graph of the Average Treatment Length per week 
ggplot(gg_avg_length_per_week_by_treatment, aes(x = Week, y = Average, shape = Treatment)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3, fill = "white", stroke = 1.2) + 
  geom_errorbar(aes(ymin = Average - SD, ymax = Average + SD),
  width = 0.2, alpha = 0.6) + 
  labs(
  title = expression("Mean length of " * italic("Gracilaria gracilis") * " by treatment and week"),
    x = "Week",
    y = "Mean length (cm)",
    color = "Treatment") +
  theme_minimal() +
    theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", size = 0.5)) +
  scale_color_brewer(palette = "Dark2") + 
   scale_shape_manual(values = c(21, 24, 22))

# Tabulating Average length
gg_avg_length_per_week_by_treatment$Date <- as.Date(gg_avg_length_per_week_by_treatment$Date) # proper format

gg_length_table <- gg_avg_length_per_week_by_treatment %>% # Getting days
  group_by(Treatment) %>%
  mutate(Days = as.numeric(Date - min(Date))) %>%
  replace_na(list(Days = 0)) %>%
  ungroup()

gg_length_table1 <- gg_length_table %>% # Joining Week and Days
  mutate(Week_Days = paste0("Week ", Week, "  ( ",Days," )")) %>%
  select(Treatment, Week_Days, Average)

gg_length_table1 <- gg_length_table1 %>% # Long to Wide format 
  pivot_wider(
    names_from = Week_Days,
    values_from = Average)

gg_tab_length <- gg_length_table1 %>%
  gt( ) %>%
  fmt_number(
    columns = everything(),
    decimals = 2
  ) %>%
  tab_header(title = md("**Total length (mean, cm)**")) %>%
  tab_spanner(
    label = "Mean length (cm)",
    columns = -Treatment) 

```

Specific Growth Rate (SGR) for Length

```{r}



#  Calculate specific growth rate (SGR) of length of both species of seaweed
gg_avg_length$Date <- as.Date(gg_avg_length$Date) # proper format

gg_rgr_length <- gg_avg_length %>%
arrange(Tank_No) %>% 
group_by(Tank_No,Treatment) %>%
  mutate(Delta_Time = as.numeric((Date) - lag(Date)),
  RGR_Length = (log(Average) - lag(log(Average))) / Delta_Time,
  Tot_RGR = (Average[Week == 3]-Average[Week == 0])/as.numeric(max(Date)-min(Date)), 
  RGR_Per =  ((log(Average) - lag(log(Average))) / Delta_Time)*100)

 rgr_weekly_sd <- gg_rgr_length %>%
  group_by(Week,Treatment) %>%
  mutate(
    SD_RGR_Per = sd(RGR_Per, na.rm = TRUE),
    N = sum(!is.na(RGR_Per)), SE_RGR_Per = SD_RGR_Per / sqrt(N)) %>%
   ungroup()

gg_avg_rgr_length <- rgr_weekly_sd %>%
  group_by(Treatment, Week) %>%
  summarise(
  RGR_Length = mean(RGR_Length,na.rm = TRUE), N = n(),
  Tot_RGR = mean(Tot_RGR, na.rm = TRUE),
  RGR_Per = mean(RGR_Per, na.rm = TRUE),
  RGR_Per_SD = mean(SD_RGR_Per, na.rm = TRUE),
  RGR_Per_SE = mean(SE_RGR_Per),
  Date = max(Date), .groups = "drop") %>%
ungroup()
 



```

SGR Figures

```{r, echo=FALSE, warning=FALSE}
    
# Bar Graph for SGR percentage of length of GG
ggplot(gg_avg_rgr_length, aes(x = factor(Week), y = RGR_Per, fill = Treatment)) +
  geom_col(position = position_dodge(width = 0.8), color = "black") +
  scale_fill_grey(start = 0.3, end = 0.8) + 
  geom_errorbar(aes(ymin = RGR_Per - RGR_Per_SE, 
  ymax = RGR_Per + RGR_Per_SE), width = 0.2, 
  position = position_dodge(width = 0.9)) + 
  scale_y_continuous(breaks = seq(-3,3,by = 0.5)) + 
   geom_hline(yintercept = 0, colour = "black", linewidth = 0.6) +
  labs(
  title = expression("Daily Growth Rate of " * italic("Gracilaria gracilis") * " by Treatment and Week"),
    x = "Week",
    y = "DGR (%)",
    fill = "Treatment"
  ) +
   theme_classic(base_size = 14) +  
   theme(
    panel.grid = element_blank(),                 
    axis.line = element_line(color = "black", size = 0.5), 
    axis.ticks = element_line(color = "black"),   
    legend.key = element_blank()                 
  )

# SGR Table
gg_rgr_length_table <- gg_avg_rgr_length %>%
  mutate(Date = as.Date(Date)) %>% 
  group_by(Treatment) %>%
  mutate(Days_cum = as.numeric(Date - first(Date))) %>%
  mutate(RGR_Days = paste0(lag(Days_cum, default = 0), "-", Days_cum)) %>%
  mutate(RGR_Length = replace_na(RGR_Length, 0)) %>%
  ungroup()

gg_rgr_length_table1 <- gg_rgr_length_table %>%
  select(Treatment, RGR_Days, RGR_Length, Tot_RGR) %>%
  pivot_wider(
    names_from = RGR_Days,
    values_from = RGR_Length
  )

rgr_order <- gg_rgr_length_table %>%
  arrange(Days_cum) %>%
  pull(RGR_Days) %>%
  unique()

gg_rgr_length_table1 <- gg_rgr_length_table %>%
  select(Treatment, RGR_Days, RGR_Length, Tot_RGR) %>%
  pivot_wider(
    names_from = RGR_Days,
    values_from = RGR_Length
  ) %>%
   select(Treatment, all_of(rgr_order), Tot_RGR)

gg_tab_rgr_length <- gg_rgr_length_table1 %>% gt() %>% tab_header(
    title = md("**Specific growth rate**"),  ) %>% 
  fmt_number(
    columns = where(is.numeric),
    decimals = 5
  ) %>% cols_label(
    Tot_RGR = "(0-21)"
  ) %>% tab_spanner(
    label = "RGR per Time Period (Days)",
    columns = c("0-0":"14-21")
  ) %>% tab_spanner(label = "Total RGR (Days)",
    columns = Tot_RGR
  ) %>%tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_stub(rows = everything())
  )

```

### Weight

Filtering the data and calculating averages

```{r}

gg_weight <- final_data %>%
  filter(Measurement_Type == "Weight", Species == "Gracilaria gracilis" ) 

# Calculate mean, SD and SE
gg_avg_weight <- gg_weight %>% 
  group_by(Treatment,Tank_No,Week) %>%
  summarise(
  Average = mean(Value,na.rm = TRUE),N = n(), SD = sd(Value, na.rm = TRUE), SE =       SD / sqrt(N),Date = max(Date),.groups = "drop") %>%
  ungroup()

# Average treatment weight per week
 gg_avg_weight_per_week_by_treatment <- gg_avg_weight %>% 
  group_by(Week,Treatment) %>%
  summarise(Average = mean(Average,na.rm = TRUE), N= n(), SD = mean(SD,na.rm = TRUE), SE = mean(SE,na.rm = TRUE), Date = max(Date), .groups = "drop") %>%
  ungroup()

```

Testing ANOVA Assumptions, creating models and running post hoc test

```{r}

# Shapiro-Wilk normality test to test normality of gg avg weight per treatment per week
shapiro.test(gg_avg_weight_per_week_by_treatment$Average) # assumption of normality met

# Levene's Test to check for homogeneity of treatments in gg avg weight per treatment per week
leveneTest(Average ~ Treatment, center = mean, data = gg_avg_weight_per_week_by_treatment) # assumption of equal variances met

# Model of average weight per treatment per week
mGAW <-  lm(Average ~ Treatment + Week, data = gg_avg_weight_per_week_by_treatment)
summary(mGAW)
anova(mGAW)

# Post hoc test
emmeans_mGAW <- emmeans(mGAW, pairwise ~ Treatment)
emmeans_mGAW$emmeans
emmeans_mGAW$contrasts
```

Figures of average weight

```{r, echo=FALSE, warning=FALSE}

# Box plot distribution of weight per treatment per week of GG
ggplot(gg_avg_weight, aes(x = factor(Week), y = Average)) +
  geom_boxplot(position = position_dodge(width = 0.6),alpha =0.5, fill = "white")+ facet_wrap(~ Treatment, scale = "free_x") +
labs(
title = expression("Mean weight of " * italic("Gracilaria gracilis") * " by treatment and    week"),
    y = "Mean weight (g)",
    x = "Week") +
  theme_bw() + 
  theme(
    panel.grid = element_blank(),           
    axis.line = element_line(color = "black") 
  )
    
# Line graph of average weight per treatment per week of GG
ggplot(gg_avg_weight_per_week_by_treatment, aes(x = Week, y = Average, shape = Treatment)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3,  fill = "white", stroke = 1.2) +
  geom_errorbar(aes(ymin = Average - SD, ymax = Average + SD),
                width = 0.1, size = 0.5) +
  labs(
title = expression("Mean weight of " * italic("Gracilaria gracilis") * " by treatment and    week"),
    x = "Week",
    y = "Mean weight (g)",
    color = "Treatment"
  ) +
  theme_minimal() +
    theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", size = 0.5)
  ) +
  scale_color_brewer(palette = "Dark2") + 
  scale_shape_manual(values = c(21, 24, 22))  

# Tables

# Average weight of GG
gg_avg_weight_per_week_by_treatment$Date <- as.Date(gg_avg_weight_per_week_by_treatment$Date) # proper format

# Average Weight Table
gg_weight_table <- gg_avg_weight_per_week_by_treatment %>% # Getting days
  group_by(Treatment) %>%
  mutate(Days = as.numeric(Date - min(Date))) %>%
  replace_na(list(Days = 0)) %>%
  ungroup()

gg_weight_table1 <- gg_weight_table %>% # Joining Week and Days
  mutate(Week_Days = paste0("Week ", Week, "  ( ",Days," )")) %>%
  select(Treatment, Week_Days, Average)

gg_weight_table1 <- gg_weight_table1 %>% # Long to Wide format 
  pivot_wider(
    names_from = Week_Days,
    values_from = Average)

gg_tab_weight <- gg_weight_table1 %>%
  gt( ) %>%
  fmt_number(
    columns = everything(),
    decimals = 2
  ) %>%
  tab_header(title = "Average Treatment Weight per Week of GG") %>%
  tab_spanner(
    label = "Wet weight (mean, g)",
    columns = -Treatment)

```

SGR

```{r}

# Calculating SGR of weight by treatment
gg_avg_weight$Date <- as.Date(gg_avg_weight$Date) # proper format

gg_rgr_weight <- gg_avg_weight %>%
arrange(Tank_No) %>% 
group_by(Tank_No,Treatment) %>%
  mutate(Delta_Time = as.numeric((Date) - lag(Date)),
  RGR_Weight = (log(Average) - lag(log(Average))) / Delta_Time,
  Tot_RGR = (Average[Week == 3]-Average[Week == 0])/as.numeric(max(Date)-min(Date)),
  RGR_Per =  ((log(Average) - lag(log(Average))) / Delta_Time)*100)

rgr_weekly_sd <- gg_rgr_weight %>%
  group_by(Week,Treatment) %>%
  mutate(
    RGR_Per_SD = sd(RGR_Per, na.rm = TRUE),
     N = sum(!is.na(RGR_Per)), SE_RGR_Per = RGR_Per_SD / sqrt(N)) %>%
  ungroup()
    
gg_avg_rgr_weight <- rgr_weekly_sd %>%
  group_by(Treatment, Week) %>%
  summarise(
  RGR_Weight = mean(RGR_Weight,na.rm = TRUE), N = n(),
  Tot_RGR = mean(Tot_RGR, na.rm = TRUE),
   RGR_Per =mean(RGR_Per, na.rm = TRUE),
   RGR_Per_SD =mean(RGR_Per_SD, na.rm = TRUE),
   RGR_Per_SE = mean(SE_RGR_Per, na.rm = TRUE),
  Date = max(Date), .groups = "drop")
```

SGR Figures

```{r, echo=FALSE, warning=FALSE}
    
# Bar graph of RGR Percentage 
ggplot(gg_avg_rgr_weight, aes(x = factor(Week), y = RGR_Per, fill = Treatment)) +
  geom_col(position = position_dodge(width = 0.8), color = "black") +
  scale_fill_grey(start = 0.3, end = 0.8) +  
  geom_errorbar(aes(ymin = RGR_Per - RGR_Per_SE, 
  ymax = RGR_Per + RGR_Per_SE), width = 0.2, 
  position = position_dodge(width = 0.9)) + 
  scale_y_continuous(breaks = seq(-3,6,by = 0.5)) + 
  geom_hline(yintercept = 0, colour = "black", linewidth = 0.6) +
  labs(
  title = expression("Daily Growth Rate of " * italic("Gracilaria gracilis") * " by Treatment and Week"),
    x = "Week",
    y = "DGR (%)",
    fill = "Treatment"
  ) +
  theme_classic(base_size = 14) +  
  theme(
    panel.grid = element_blank(),                
    axis.line = element_line(color = "black", size = 0.5), 
    axis.ticks = element_line(color = "black"),   
    legend.key = element_blank()                  
  )


# SGR weight table GG
gg_rgr_weight_table <- gg_avg_rgr_weight %>%
  mutate(Date = as.Date(Date)) %>% 
  group_by(Treatment) %>%
  mutate(Days_cum = as.numeric(Date - first(Date))) %>%
  mutate(RGR_Days = paste0(lag(Days_cum, default = 0), "-", Days_cum)) %>%
  mutate(RGR_Weight = replace_na(RGR_Weight, 0)) %>%             
  ungroup()

gg_rgr_weight_table1 <- gg_rgr_weight_table %>%
  select(Treatment, RGR_Days, RGR_Weight, Tot_RGR) %>%
  pivot_wider(
    names_from = RGR_Days,
    values_from = RGR_Weight
  )

rgr_order <- gg_rgr_weight_table %>%
  arrange(Days_cum) %>%
  pull(RGR_Days) %>%
  unique()

gg_rgr_weight_table1 <- gg_rgr_weight_table %>%
  select(Treatment, RGR_Days, RGR_Weight, Tot_RGR) %>%
  pivot_wider(
    names_from = RGR_Days,
    values_from = RGR_Weight
  ) %>%
   select(Treatment, all_of(rgr_order), Tot_RGR)

gg_tab_rgr_weight <- gg_rgr_weight_table1 %>% 
  gt() %>% tab_header(
    title = md("**Relative Growth Rate (Weight)**"),
    subtitle = "Per time period with total RGR at right"
  ) %>% fmt_number(
    columns = where(is.numeric),
    decimals = 5
  ) %>% cols_label(
    Tot_RGR = "(0-21)"
  ) %>% tab_spanner(
    label = "RGR per Time Period (Days)",
    columns = c("0-0":"14-21")
  ) %>% tab_spanner(label = "Total SGR (Days)",
    columns = Tot_RGR
  ) %>%tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_stub(rows = everything())
  )
  
```

## *Gracilariopsis longissima*

### Length

Filtering the data and calculating averages

```{r}

gl_length <- final_data %>%
  filter(Measurement_Type == "Length", Species == "Gracilariopsis longissima" ) 

# Calculate mean SD and SE
gl_avg_length <- gl_length %>%
  group_by(Treatment, Tank_No,Week) %>%
  summarise(Average = mean(Value,na.rm = TRUE),N = n(), SD = sd(Value), 
            SE = SD / sqrt(N),Date = max(Date),.groups = "drop") %>% 
  ungroup() 

# Average treatment length per week
gl_avg_length_per_week_by_treatment <- gl_avg_length  %>%
  group_by( Week,Treatment) %>%
  summarise(
    Average = mean(Average,na.rm = TRUE), N = n(),
    SD = mean(SD,na.rm = TRUE), SE = mean(SE,na.rm = TRUE),
    Date = max(Date), .groups = "drop")

```

Testing ANOVA Assumptions, modelling and post hoc test

```{r}

# Shapiro-Wilk normality test 
shapiro.test(gl_avg_length_per_week_by_treatment$Average) # normality assumption not met

# Levene's Test to check for homogeneity of treatments 
leveneTest(Average ~ Treatment, center = mean, data = gl_avg_length_per_week_by_treatment) # equal variances assumption not met
       
# Use generalised least squares model instead of linear model
m_LAL <- gls(
  Average ~ Treatment + Week,
  weights = varIdent(form = ~1 | Treatment),  
  data = gl_avg_length_per_week_by_treatment
)
summary(m_LAL)
anova(m_LAL)

# Post hoc test
emmeans_gls <- emmeans(m_LAL, pairwise ~ Treatment)
emmeans_gls$emmeans
emmeans_gls$contrasts

```

Figures of average length

```{r, echo=FALSE}

# Line graph of average length per treatment per week 
ggplot(gl_avg_length_per_week_by_treatment, aes(x = Week, y = Average, shape = Treatment)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3, fill = "white", stroke = 1.2) + 
  geom_errorbar(aes(ymin = Average - SD, ymax = Average + SD), width = 0.2, alpha = 0.6) +
  labs(
title = expression("Mean length of " * italic("Gracilariopsis longissima ") * " by treatment and week"),

    x = "Week",
    y = "Mean Length (cm)",
    color = "Treatment"
  ) +
  theme_minimal() +
    theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black", size = 0.5)  ) +
  scale_color_brewer(palette = "Dark2") + 
  scale_shape_manual(values = c(21, 24, 22))  

# Average length Table

gl_avg_length_per_week_by_treatment$Date <- as.Date(gl_avg_length_per_week_by_treatment$Date) # proper format

gl_length_table <- gl_avg_length_per_week_by_treatment %>% 
  group_by(Treatment) %>%
  mutate(Days = as.numeric(Date - min(Date))) %>%
  replace_na(list(Days = 0)) %>%
  ungroup()

gl_length_table1 <- gl_length_table %>% 
  mutate(Week_Days = paste0("Week ", Week, "  ( ",Days," )")) %>%
  select(Treatment, Week_Days, Average)

gl_length_table1 <- gl_length_table1 %>% 
  pivot_wider(
    names_from = Week_Days,
    values_from = Average)

gl_tab_length <- gl_length_table1 %>%
  gt( ) %>%
  fmt_number(
    columns = everything(),
    decimals = 2
  ) %>%
    tab_header(title = md("**Total length (mean, cm)**")) %>%
  tab_spanner(
    label = " Mean length (cm)",
    columns = -Treatment)

```

SGR (Length)

```{r}

#  Calculate specific growth rate (SGR) of length of Gracilariopsis longissima
gl_avg_length$Date <- as.Date(gl_avg_length$Date) # proper format

gl_rgr_length <- gl_avg_length %>%
arrange(Tank_No) %>% 
group_by(Tank_No,Treatment) %>%
  mutate(Delta_Time = as.numeric((Date) - lag(Date)),
  RGR_Length = (log(Average) - lag(log(Average))) / Delta_Time,
  Tot_RGR = (Average[Week == 3]-Average[Week == 0])/as.numeric(max(Date)-min(Date)),
   RGR_Per = ((log(Average) - lag(log(Average))) / Delta_Time)*100)

rgr_weekly_sd <- gl_rgr_length %>%
  group_by(Week,Treatment) %>%
  mutate(
    RGR_Per_SD = sd(RGR_Per, na.rm = TRUE),
  N = sum(!is.na(RGR_Per)), 
  SE_RGR_Per = RGR_Per_SD / sqrt(N)) %>%
  ungroup()

    
gl_avg_rgr_length <- rgr_weekly_sd %>%
  group_by(Treatment, Week) %>%
  summarise(
  RGR_Length = mean(RGR_Length,na.rm = TRUE), N = n(),
  Tot_RGR = mean(Tot_RGR, na.rm = TRUE),
    RGR_Per = mean(RGR_Per, na.rm = TRUE),
    RGR_Per_SD = mean(RGR_Per_SD, na.rm = TRUE),
  RGR_Per_SE = mean(SE_RGR_Per, na.rm = TRUE),
  Date = max(Date), .groups = "drop") %>%
ungroup()

```

SGR Figures

```{r, echo=FALSE, warning=FALSE}
    
# Bar Graph for SGR percentage of length 
ggplot(gl_avg_rgr_length, aes(x = factor(Week), y = RGR_Per, fill = Treatment)) +
  geom_col(position = position_dodge(width = 0.8), color = "black") +
  geom_errorbar(aes(ymin = RGR_Per - RGR_Per_SE, 
  ymax = RGR_Per + RGR_Per_SE), width = 0.2, 
  position = position_dodge(width = 0.9)) + 
  scale_y_continuous(breaks = seq(-3,6,by = 0.5)) + 
  geom_hline(yintercept = 0, colour = "black", linewidth = 0.6) +
  scale_fill_grey(start = 0.3, end = 0.8) + 
  labs(
title = expression("Daily Growth Rate of " * italic("Gracilariopsis longissima") * " by Treatment and Week"),
    x = "Week",
    y = "DGR (%)",
    fill = "Treatment"
  ) +
  theme_minimal(base_size = 14) +
 theme(
    panel.grid = element_blank(),                 
    axis.line = element_line(color = "black", size = 0.5), 
    axis.ticks = element_line(color = "black"),   
    legend.key = element_blank()                  
  )

# SGR Length Table
gl_rgr_length_table <- gl_avg_rgr_length %>%
  mutate(Date = as.Date(Date)) %>% 
  group_by(Treatment) %>%
  mutate(Days_cum = as.numeric(Date - first(Date))) %>%
  mutate(RGR_Days = paste0(lag(Days_cum, default = 0), "-", Days_cum)) %>%
  mutate(RGR_Length = replace_na(RGR_Length, 0)) %>%
  ungroup()

gl_rgr_length_table1 <- gl_rgr_length_table %>%
  select(Treatment, RGR_Days, RGR_Length, Tot_RGR) %>%
  pivot_wider(
    names_from = RGR_Days,
    values_from = RGR_Length
  )

rgr_order <- gl_rgr_length_table %>%
  arrange(Days_cum) %>%
  pull(RGR_Days) %>%
  unique()

gl_rgr_length_table1 <- gl_rgr_length_table %>%
  select(Treatment, RGR_Days, RGR_Length, Tot_RGR) %>%
  pivot_wider(
    names_from = RGR_Days,
    values_from = RGR_Length
  ) %>%
   select(Treatment, all_of(rgr_order), Tot_RGR)

gl_tab_rgr_length <- gl_rgr_length_table1 %>% 
  gt() %>% tab_header(
    title = md("**Specific growth rate**"),
  ) %>% fmt_number(
    columns = where(is.numeric),
    decimals = 5
  ) %>% cols_label(
    Tot_RGR = "(0-20)"
  ) %>% tab_spanner(
    label = "SGR per Time Period (Days)",
    columns = c("0-0":"13-20")
  ) %>% tab_spanner(label = "Total RGR (Days)",
    columns = Tot_RGR
  ) %>%tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_stub(rows = everything())
  )

```

### Weight

```{r}

#  Specific growth rate (SGR) of weight of Gracilariopsis longissima
gl_weight <- final_data %>%
  filter(Measurement_Type == "Weight", Species == "Gracilariopsis longissima" ) 

# Calculate mean, SD and SE
gl_avg_weight <- gl_weight %>%
  group_by(Treatment,Tank_No,Week) %>%
  summarise(
  Average = mean(Value,na.rm = TRUE),N = n(), SD = sd(Value, na.rm = TRUE), SE = SD / sqrt(N),Date = max(Date),.groups = "drop") %>%
  ungroup()

# Average treatment weight per week
gl_avg_weight_per_week_by_treatment <- gl_avg_weight %>%
  group_by(Week,Treatment) %>%
  summarise(Average = mean(Average,na.rm = TRUE), N= n(), SD = mean(SD,na.rm = TRUE), SE = mean(SE,na.rm = TRUE), Date = max(Date), .groups = "drop") %>%
  ungroup()

```

Testing ANOVA assumptions and modelling

```{r}

# Shapiro-Wilk normality test 
shapiro.test(gl_avg_weight_per_week_by_treatment$Average) # normality assumption met

# Levene's Test to check for homogeneity 
leveneTest(Average ~ Treatment, center = mean, data = gl_avg_weight_per_week_by_treatment) # equal variances assumption met

# Model average weight per treatment per week
mLAW <-  lm(Average ~ Treatment + Week, data = gl_avg_weight_per_week_by_treatment)

summary(mLAW)
anova(mLAW)
aov(mLAW)

# Post hoc test
emmeans_mLAW <- emmeans(mLAW, pairwise ~ Treatment)
emmeans_mLAW$emmeans
emmeans_mLAW$contrasts

```

Figures average weight

```{r, echo=FALSE, warning=FALSE}

# Box plot of weight of GL
ggplot(gl_avg_weight, aes(x = factor(Week), y = Average)) +
  geom_boxplot(position = position_dodge(width = 0.6),alpha =0.5, fill = "white")+ facet_wrap(~ Treatment, scale = "free_x") +
labs(
  title = expression("Mean weight of " * italic("Gracilariopsis longissima") * " by treatment and week"),
    y = "Mean weight (g)",
    x = "Week") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),                   
    axis.line = element_line(color = "black"), 
    panel.border = element_rect(color = "black", fill = NA, size = 1), 
    strip.background = element_rect(color = "black", fill = "white", size = 1), 
    strip.text = element_text(face = "bold")
  )
    
# Line graph of average weight per treatment per week
ggplot(gl_avg_weight_per_week_by_treatment, aes(x = Week, y = Average, shape = Treatment)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3, , fill = "white", stroke = 1.2) +  
  geom_errorbar(aes(ymin = Average - SD, ymax = Average + SD),
                width = 0.1, alpha = 0.6) +  
  labs(
    title = expression("Mean weight of " * italic("Gracilariopsis longissima") * " by treatment and    week"),
    x = "Week",
    y = "Mean Weight (g)",
    color = "Treatment"
  ) +
  theme_minimal() +
    theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", size = 0.5)  ) +
  scale_color_brewer(palette = "Dark2") + 
  scale_shape_manual(values = c(21, 24, 22))

# Table GL average weight
gl_avg_weight_per_week_by_treatment$Date <- as.Date(gl_avg_weight_per_week_by_treatment$Date) # proper format

gl_weight_table <- gl_avg_weight_per_week_by_treatment %>% # Getting days
  group_by(Treatment) %>%
  mutate(Days = as.numeric(Date - min(Date))) %>%
  replace_na(list(Days = 0)) %>%
  ungroup()

gl_weight_table1 <- gl_weight_table %>% # Joining Week and Days
  mutate(Week_Days = paste0("Week ", Week, "  ( ",Days," )")) %>%
  select(Treatment, Week_Days, Average)

gl_weight_table1 <- gl_weight_table1 %>% # Long to Wide format 
  pivot_wider(
    names_from = Week_Days,
    values_from = Average)

gl_tab_weight <- gl_weight_table1 %>%
  gt( ) %>%
  fmt_number(
    columns = everything(),
    decimals = 2
  ) %>%
  tab_header(title = md("**Average Treatment Weight per Week of GL**")) %>%
  tab_spanner(
    label = "Wet weight (mean, g)",
    columns = -Treatment)

```

SGR

```{r}

gl_avg_weight$Date <- as.Date(gl_avg_weight$Date) # make sure date is formatted

gl_rgr_weight <- gl_avg_weight %>%
arrange(Tank_No) %>% 
group_by(Tank_No,Treatment) %>%
  mutate(Delta_Time = as.numeric((Date) - lag(Date)),
  RGR_Weight = (log(Average) - lag(log(Average))) / Delta_Time,
  Tot_RGR = (Average[Week == 3]-Average[Week == 0])/as.numeric(max(Date)-min(Date)),
  RGR_Per = ((log(Average) - lag(log(Average))) / Delta_Time)*100)

rgr_weekly_sd <- gl_rgr_weight %>%
  group_by(Week,Treatment) %>%
  mutate(
    RGR_Per_SD = sd(RGR_Per, na.rm = TRUE),
  N = sum(!is.na(RGR_Per)), 
  SE_RGR_Per = RGR_Per_SD / sqrt(N)) %>%
  ungroup()

    
gl_avg_rgr_weight <- rgr_weekly_sd %>%
  group_by(Treatment, Week) %>%
  summarise(
  RGR_Weight = mean(RGR_Weight,na.rm = TRUE), N = n(),
  Tot_RGR = mean(Tot_RGR, na.rm = TRUE),
  RGR_Per = mean(RGR_Per, na.rm = TRUE),
  RGR_Per_SD = mean(RGR_Per_SD, na.rm = TRUE),
  RGR_Per_SE = mean(SE_RGR_Per, na.rm = TRUE),
  Date = max(Date), .groups = "drop") %>%
ungroup()

```

SGR Figures

```{r, echo=FALSE, warning=FALSE}
    
# Bar graph of SGR Weight of GL
 ggplot(gl_avg_rgr_weight, aes(x = factor(Week), y = RGR_Per, fill = Treatment)) +
  geom_col(position = position_dodge(width = 0.8), color = "black") +
  geom_errorbar(aes(ymin = RGR_Per - RGR_Per_SE, 
  ymax = RGR_Per + RGR_Per_SE), width = 0.2, 
  position = position_dodge(width = 0.9)) + 
  scale_y_continuous(breaks = seq(-3,6,by = 0.5)) + 
  geom_hline(yintercept = 0, colour = "black", linewidth = 0.6) +
  scale_fill_grey(start = 0.3, end = 0.8) + 
  labs(
  title = expression("Daily Growth Rate of " * italic("Gracilariopsis longissima") * " by Treatment and Week"),
    x = "Week",
    y = "DGR (%)",
    fill = "Treatment"
  ) +
  theme_minimal(base_size = 14) +
 theme(
    panel.grid = element_blank(),               
    axis.line = element_line(color = "black", size = 0.5),
    axis.ticks = element_line(color = "black"),   
    legend.key = element_blank()                  
  )

# Table of SGR Weight 
gl_rgr_weight_table <- gl_avg_rgr_weight %>%
  mutate(Date = as.Date(Date)) %>% 
  group_by(Treatment) %>%
  mutate(Days_cum = as.numeric(Date - first(Date))) %>%
  mutate(RGR_Days = paste0(lag(Days_cum, default = 0), "-", Days_cum)) %>%
  mutate(RGR_Weight = replace_na(RGR_Weight, 0)) %>%             
  ungroup()

gl_rgr_weight_table1 <- gl_rgr_weight_table %>%
  select(Treatment, RGR_Days, RGR_Weight, Tot_RGR) %>%
  pivot_wider(
    names_from = RGR_Days,
    values_from = RGR_Weight
  )

rgr_order <- gl_rgr_weight_table %>%
  arrange(Days_cum) %>%
  pull(RGR_Days) %>%
  unique()

gl_rgr_weight_table1 <- gl_rgr_weight_table %>%
  select(Treatment, RGR_Days, RGR_Weight, Tot_RGR) %>%
  pivot_wider(
    names_from = RGR_Days,
    values_from = RGR_Weight
  ) %>%
   select(Treatment, all_of(rgr_order), Tot_RGR)

gl_tab_rgr_weight <- gl_rgr_weight_table1 %>% 
  gt() %>% tab_header(
    title = md("**Relative Growth Rate (Weight)**"),
    subtitle = "Per time period with total RGR at right"
  ) %>% fmt_number(
    columns = where(is.numeric),
    decimals = 5
  ) %>% cols_label(
    Tot_RGR = "(0-21)"
  ) %>% tab_spanner(
    label = "SGR per Time Period (Days)",
    columns = c("0-0":"13-20")
  ) %>% tab_spanner(label = "Total SGR (Days)",
    columns = Tot_RGR
  ) %>%tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_stub(rows = everything())
  )
  

```

Relationship between total length and wet weight of three cuttings in 
*Gracilaria gracilis*

```{r, echo=FALSE, warning=FALSE}

# Scatterplots of Length VS Weight of Gracilaria gracilis (GG)

# Linking GG Length and Weight to each tank
gg_al <- gg_avg_length %>%   # extracting avg length
arrange(Tank_No) %>% 
summarise(Treatment = Treatment, AL = Average, Row_ID = row_number(), .groups = "drop")

gg_aw <- gg_avg_weight %>%    # extracting avg weight
arrange(Tank_No) %>% 
summarise(AW = Average,Row_ID = row_number(), .groups = "drop")


# Scatter plot for 10 cm treatment of GG 
gg_avg_merged <- merge(gg_al,gg_aw, by = "Row_ID" ) %>% # merging by 10
filter(Treatment == "10") 

ggplot(gg_avg_merged, aes(x = AL, y = AW)) +
  geom_point(shape = 2, size = 3, aes(color = Treatment)) + 
  geom_smooth(method = "lm", se = FALSE, color = "black") + 
  stat_poly_eq(
    aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
    formula = y ~ x,
    parse = TRUE,
    size = 4
  ) +
  labs(
    x = "Total length (cm)",
    y = "Wet weight (g)",
    color = ""  
  ) +
  scale_color_manual(values = c("black")) + 
  theme_bw(base_size = 12) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    ,
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
 
# Scatter plot for 5 cm of GG 
gg_avg_merged <- merge(gg_al,gg_aw, by = "Row_ID" ) %>% # merging by 5
filter(Treatment == "5") 

ggplot(gg_avg_merged, aes(x = AL, y = AW)) +
  geom_point(shape = 0, size = 3, aes(color = Treatment)) +  
  geom_smooth(method = "lm", se = FALSE, color = "black") + 
  stat_poly_eq(
    aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
    formula = y ~ x,
    parse = TRUE,
    size = 4
  ) +
  labs(
    x = "Total length (cm)",
    y = "Wet weight (g)",
    color = ""  
  ) +
  scale_color_manual(values = c("black")) + 
  theme_bw(base_size = 12) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    ,
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

# Scatter plot for Control of GG 
gg_avg_merged <- merge(gg_al,gg_aw, by = "Row_ID" ) %>% # merging by control
filter(Treatment == "Control") 

ggplot(gg_avg_merged, aes(x = AL, y = AW)) +
  geom_point(shape = 1, size = 4, aes(color = Treatment)) + 
  geom_smooth(method = "lm", se = FALSE, color = "black") + 
  stat_poly_eq(
    aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
    formula = y ~ x,
    parse = TRUE,
    size = 4
  ) +
  labs(
    x = "Total length (cm)",
    y = "Wet weight (g)",
    color = ""
  ) +
  scale_color_manual(values = c("black")) +
  theme_bw(base_size = 12) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    ,
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

```

Relationship between total length and wet weight of three cuttings in 
*Gracilariopsis longissima*

```{r, echo=FALSE, warning=FALSE}
# Scatterplots of Length VS Weight of Gracilariopsis longissima (GL)

# Linking GL Length and Weight to each tank

gl_al <- gl_avg_length %>%   # extracting avg length
arrange(Tank_No) %>% 
summarise(Treatment = Treatment, AL = Average, Row_ID = row_number(), .groups = "drop")

gl_aw <- gl_avg_weight %>%    # extracting avg weight
arrange(Tank_No) %>% 
summarise(AW = Average,Row_ID = row_number(), .groups = "drop")


# Scatter plot for 10 cm of GL 
gl_avg_merged <- merge(gl_al,gl_aw, by = "Row_ID" ) %>% # merging
filter(Treatment == "10") 


ggplot(gl_avg_merged, aes(x = AL, y = AW)) +
  geom_point(shape = 2, size = 3, aes(color = Treatment)) +  # Open circles
  geom_smooth(method = "lm", se = FALSE, color = "black") + 
  stat_poly_eq(
    aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
    formula = y ~ x,
    parse = TRUE,
    size = 4
  ) +
  labs(
    x = "Total length (cm)",
    y = "Wet weight (g)",
    color = ""   # empty legend title
  ) +
  scale_color_manual(values = c("black")) +  # or pick your colors
  theme_bw(base_size = 12) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    ,
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
  
# Scatter plot for 5 cm of GL  
  gl_avg_merged <- merge(gl_al,gl_aw, by = "Row_ID" ) %>% 
filter(Treatment == "5") 

ggplot(gl_avg_merged, aes(x = AL, y = AW)) +
  geom_point(shape = 0, size = 3, aes(color = Treatment)) +  # Open circles
  geom_smooth(method = "lm", se = FALSE, color = "black") + 
  stat_poly_eq(
    aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
    formula = y ~ x,
    parse = TRUE,
    size = 4
  ) +
  labs(
    x = "Total length (cm)",
    y = "Wet weight (g)",
    color = ""   # empty legend title
  ) +
  scale_color_manual(values = c("black")) +  # or pick your colors
  theme_bw(base_size = 12) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    ,
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

# Scatter plot for Control of GL
gl_avg_merged <- merge(gl_al,gl_aw, by = "Row_ID" ) %>% # merges the 2 avgs of gl
filter(Treatment == "Control") # used to plot different treatments

ggplot(gl_avg_merged, aes(x = AL, y = AW)) +
  geom_point(shape = 1, size = 4, aes(color = Treatment)) +
  geom_smooth(
    method = "lm", se = FALSE, color = "black") + 
    stat_poly_eq(
    aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
    formula = y ~ x,
    parse = TRUE,
    size = 4
  ) +
  labs(
    x = "Total length (cm)",
    y = "Wet weight (g)",
    color = ""   # empty legend title
  ) +
  scale_color_manual(values = c("black")) +  # or pick your colors
  theme_bw(base_size = 12) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    ,
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

```

Tables: ANOVA, Average lengths and Specific growth rates

```{r, echo=FALSE}

# Anova Table for GG

gg_all_merged <- merge(gg_al,gg_aw, by = "Row_ID" ) 

mcon <- aov(AL ~ AW, data = subset(gg_all_merged, Treatment == "Control"))
m5 <- aov(AL ~ AW, data = subset(gg_all_merged, Treatment == "5"))
m10 <- aov(AL ~ AW, data = subset(gg_all_merged, Treatment == "10"))


tidy_anova <- function(model, treatment_name) {
  a <- summary(model)[[1]]
  
  reg <- a[1, , drop = FALSE]
  reg_df <- data.frame(
    Treatment = treatment_name,
    Source = "Regression",
    `Sum of squares` = reg[["Sum Sq"]],
    df = reg[["Df"]],
    `Mean squares` = reg[["Mean Sq"]],
    `F value` = reg[["F value"]],
    check.names = FALSE
  )
  
  res <- a["Residuals", , drop = FALSE]
  res_df <- data.frame(
    Treatment = treatment_name,
    Source = "Residual",
    `Sum of squares` = res[["Sum Sq"]],
    df = res[["Df"]],
    `Mean squares` = res[["Mean Sq"]],
    `F value` = NA,
    check.names = FALSE
  )
  
  total_df <- data.frame(
    Treatment = treatment_name,
    Source = "Total",
    `Sum of squares` = reg[["Sum Sq"]] + res[["Sum Sq"]],
    df = reg[["Df"]] + res[["Df"]],
    `Mean squares` = NA,
    `F value` = NA,
    check.names = FALSE
  )
  
  rbind(reg_df, res_df, total_df)
}

# Critical F function

get_dfs <- function(model) {
  a <- summary(model)[[1]]
  df1 <- a[1, "Df"] 
  df2 <- a["Residuals", "Df"] 
  c(df1 = df1, df2 = df2)
}

# Critical F table

critical_Fs <- tibble(
  Treatment = c("Control", "5", "10"),
  df1 = c(get_dfs(mcon)["df1"], get_dfs(m5)["df1"], get_dfs(m10)["df1"]),
  df2 = c(get_dfs(mcon)["df2"], get_dfs(m5)["df2"], get_dfs(m10)["df2"])
) %>%
  mutate(F0.05 = qf(0.95, df1, df2)) %>%
  select(Treatment, F0.05)


# Create Anova Table

anova_df <- bind_rows(
  tidy_anova(mcon, "Control"),
  tidy_anova(m5, "5"),
  tidy_anova(m10, "10")
) %>%
  left_join(critical_Fs, by = "Treatment") %>%
  rename(`Cutting length` = Treatment) %>%
  group_by(`Cutting length`) %>%
  mutate(`Cutting length` = ifelse(row_number() == 1, `Cutting length`, "")) %>%
  ungroup()

anova_df %>%
  gt() %>%
  fmt_number(
    columns = where(is.numeric),
    decimals = 2
  ) %>%
  tab_header(
        title = md("**Analysis of variance between total length and wet weight of three kinds
               of cuttings in _Gracilaria gracilis_**")  ) %>%
  tab_options(
    table.font.size = px(14),
    data_row.padding = px(5),
    table_body.hlines.style = "none" 
  ) %>%
  tab_style( 
    style = cell_borders(sides = "bottom", color = "black", weight = px(1)),
    locations = cells_body(rows = Source == "Total")
  )


# Anova Table for GL

gl_all_merged <- merge(gl_al,gl_aw, by = "Row_ID" ) 

mcon <- aov(AL ~ AW, data = subset(gl_all_merged, Treatment == "Control"))
m5 <- aov(AL ~ AW, data = subset(gl_all_merged, Treatment == "5"))
m10 <- aov(AL ~ AW, data = subset(gl_all_merged, Treatment == "10"))


tidy_anova <- function(model, treatment_name) {
  a <- summary(model)[[1]]
  
  reg <- a[1, , drop = FALSE]
  reg_df <- data.frame(
    Treatment = treatment_name,
    Source = "Regression",
    `Sum of squares` = reg[["Sum Sq"]],
    df = reg[["Df"]],
    `Mean squares` = reg[["Mean Sq"]],
    `F value` = reg[["F value"]],
    check.names = FALSE
  )
  
  res <- a["Residuals", , drop = FALSE]
  res_df <- data.frame(
    Treatment = treatment_name,
    Source = "Residual",
    `Sum of squares` = res[["Sum Sq"]],
    df = res[["Df"]],
    `Mean squares` = res[["Mean Sq"]],
    `F value` = NA,
    check.names = FALSE
  )
  
  total_df <- data.frame(
    Treatment = treatment_name,
    Source = "Total",
    `Sum of squares` = reg[["Sum Sq"]] + res[["Sum Sq"]],
    df = reg[["Df"]] + res[["Df"]],
    `Mean squares` = NA,
    `F value` = NA,
    check.names = FALSE
  )
  
  rbind(reg_df, res_df, total_df)
}

# Critical F function

get_dfs <- function(model) {
  a <- summary(model)[[1]]
  df1 <- a[1, "Df"] 
  df2 <- a["Residuals", "Df"] 
  c(df1 = df1, df2 = df2)
}

# Critical F table

critical_Fs <- tibble(
  Treatment = c("Control", "5", "10"),
  df1 = c(get_dfs(mcon)["df1"], get_dfs(m5)["df1"], get_dfs(m10)["df1"]),
  df2 = c(get_dfs(mcon)["df2"], get_dfs(m5)["df2"], get_dfs(m10)["df2"])
) %>%
  mutate(F0.05 = qf(0.95, df1, df2)) %>%
  select(Treatment, F0.05)


# Create Anova Table

anova_df <- bind_rows(
  tidy_anova(mcon, "Control"),
  tidy_anova(m5, "5"),
  tidy_anova(m10, "10")
) %>%
  left_join(critical_Fs, by = "Treatment") %>%
  rename(`Cutting length` = Treatment) %>%
  group_by(`Cutting length`) %>%
  mutate(`Cutting length` = ifelse(row_number() == 1, `Cutting length`, "")) %>%
  ungroup()

anova_df %>%
  gt() %>%
  fmt_number(
    columns = where(is.numeric),
    decimals = 2
  ) %>%
  tab_header(
    title = md("**Analysis of variance between total length and wet weight of three kinds
               of cuttings in _Gracilariopsis longissima_**")) %>%
  tab_options(
    table.font.size = px(14),
    data_row.padding = px(5),
    table_body.hlines.style = "none" 
  ) %>%
  tab_style( 
    style = cell_borders(sides = "bottom", color = "black", weight = px(1)),
    locations = cells_body(rows = Source == "Total")
  )

```

Total length (cm) of cuttings cut from individuals of *Gracilaria gracilis*

```{r, echo=FALSE}

# Join GG Length and SGR table
browsable(tagList( gg_tab_length,gg_tab_rgr_length))

```

Total length (cm) of cuttings cut from individuals of *Gracilariopsis longissima*

```{r, echo=FALSE}

# Join GL Length and SGR table

browsable(tagList( gl_tab_length,gl_tab_rgr_length))

```

Wet weight (g) of cuttings cut from individuals of *Gracilaria gracilis*

```{r, echo=FALSE}

# Join GG Weight and SGR table
browsable(tagList( gg_tab_weight,gg_tab_rgr_weight))

```

Wet weight (g) of cuttings cut from individuals of *Gracilariopsis longissima*

```{r, echo=FALSE}

# Join GL Weight and SGR table
browsable(tagList( gl_tab_weight,gl_tab_rgr_weight))

```


Combine Weight and Length DGRs in bar graph for *Gracilaria gracilis*

```{r}

library(ggplot2)
library(dplyr)

# Combine the two datasets, adding a column to distinguish length vs weight
gg_rgr_combined <- bind_rows(
  gg_avg_rgr_length %>% mutate(Type = "Length"),
  gg_avg_rgr_weight %>% mutate(Type = "Weight")
)


gg_rgr_combined_plot <- gg_rgr_combined %>%
  filter(Treatment %in% c(5, 10)) %>%
  group_by(Treatment, Type) %>%
  summarise(Mean_RGR_Per = mean(RGR_Per, na.rm = TRUE), 
            RGR_Per_SE = mean(RGR_Per_SE, na.rm = TRUE),.groups = "drop") %>%
  ungroup()

# Plot
ggplot(gg_rgr_combined_plot, aes(x = Treatment, y = Mean_RGR_Per, fill = Type)) +
  geom_col(position = position_dodge(width = 0.3), color = "black", width = 0.2) +
    geom_errorbar(aes(ymin = Mean_RGR_Per, ymax = Mean_RGR_Per + RGR_Per_SE),
                width = 0.2, position = position_dodge(width = 0.3)) +
  scale_fill_grey(start = 1, end = 0) +
  geom_smooth(aes(group = Type), color = "black",method = "lm", se = FALSE, linetype = "solid") +
  labs(
    title = expression("Daily Growth Rate of " * italic("Gracilaria gracilis") * " by Treatment"),
    x = "Treatment",
    y = "DGR (%)",
    fill = "",
    color = ""
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.line = element_line(color = "black", size = 0.5),
    axis.ticks = element_line(color = "black"),
    legend.position = "top"
  )
```

Combine Weight and Length DGRs in bar graph for *Gracilariopsis longissima*

```{r}



# Combine the two datasets, adding a column to distinguish length vs weight
gl_rgr_combined <- bind_rows(
  gl_avg_rgr_length %>% mutate(Type = "Length"),
  gl_avg_rgr_weight %>% mutate(Type = "Weight")
)




gl_rgr_combined_plot <- gl_rgr_combined %>%
  filter(Treatment %in% c(5, 10)) %>%
  group_by(Treatment, Type) %>%
  summarise(Mean_RGR_Per = mean(RGR_Per, na.rm = TRUE), 
            RGR_Per_SE = mean(RGR_Per_SE, na.rm = TRUE),.groups = "drop") %>%
  ungroup()

# Plot
ggplot(gl_rgr_combined_plot, aes(x = Treatment, y = Mean_RGR_Per, fill = Type)) +
  geom_col(position = position_dodge(width = 0.3), color = "black", width = 0.2) +
    geom_errorbar(aes(ymin = Mean_RGR_Per, ymax = Mean_RGR_Per + RGR_Per_SE),
                width = 0.2, position = position_dodge(width = 0.3)) +
  scale_fill_grey(start = 1, end = 0) +
  geom_smooth(aes(group = Type),color = "black", method = "lm", se = FALSE, linetype = "solid") +
  labs(
    title = expression("Daily Growth Rate of " * italic("Gracilariopsis longissima") * " by Treatment"),
    x = "Treatment",
    y = "DGR (%)",
    fill = "",
    color = ""
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.line = element_line(color = "black", size = 0.5),
    axis.ticks = element_line(color = "black"),
    legend.position = "top"
  )
```

Bar graph of Final lengths and weights of *Gracilariopsis longissima*

```{r}

gl_combined <- bind_rows(
  gl_avg_length %>% mutate(Type = "Length"),
  gl_avg_weight %>% mutate(Type = "Weight"))

gl_combined_plot <- gl_combined %>%
  filter(Treatment %in% c(5, 10), Week == "3") %>%
  group_by(Treatment, Type) %>%
  summarise(Net_Yield = mean(Average, na.rm = TRUE),
   SE = mean(SE, na.rm = TRUE), .groups = "drop") %>%
  ungroup()


# plot

# Rescale length to match weight scale
# Find max values
max_weight <- max(gl_combined_plot$Net_Yield[gl_combined_plot$Type == "Weight"], na.rm = TRUE)
max_length <- max(gl_combined_plot$Net_Yield[gl_combined_plot$Type == "Length"], na.rm = TRUE)

scale_factor <- max_weight / max_length

gl_combined_plot <- gl_combined_plot %>%
  mutate(Net_Yield_scaled = ifelse(Type == "Length", Net_Yield * scale_factor, Net_Yield),
         SE_scaled = ifelse(Type == "Length", SE * scale_factor, SE))

ggplot(gl_combined_plot, aes(x = factor(Treatment), fill = Type)) +
  geom_col(aes(y = Net_Yield_scaled), 
           position = position_dodge(width = 0.4), 
           color = "black", width = 0.3) +
  geom_errorbar(aes(ymin = Net_Yield_scaled, ymax = Net_Yield_scaled + SE_scaled),
                width = 0.2, position = position_dodge(width = 0.4)) +
  scale_fill_grey(start = 1, end = 0) +
  scale_y_continuous(
    name = "Net yield (g)", 
    sec.axis = sec_axis(~ . / scale_factor, name = "Net yield (cm)")
  ) +
  labs(
    title = expression("Net yield of " * italic("Gracilariopsis longissima") * " by Treatment"),
    x = "Initial total length",
    fill = ""
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.line = element_line(color = "black", size = 0.5),
    axis.ticks = element_line(color = "black"),
    legend.position = "top"
  )


```

Bar graph of Final lengths and weights of *Gracilaria gracilis*

```{r}

gg_combined <- bind_rows(
  gg_avg_length %>% mutate(Type = "Length"),
  gg_avg_weight %>% mutate(Type = "Weight"))

gg_combined_plot <- gg_combined %>%
  filter(Treatment %in% c(5, 10), Week == "3") %>%
  group_by(Treatment, Type) %>%
  summarise(Net_Yield = mean(Average, na.rm = TRUE),
   SE = mean(SE, na.rm = TRUE))


# plot

# Rescale length to match weight scale
# Find max values
max_weight <- max(gg_combined_plot$Net_Yield[gg_combined_plot$Type == "Weight"], na.rm = TRUE)
max_length <- max(gg_combined_plot$Net_Yield[gg_combined_plot$Type == "Length"], na.rm = TRUE)

scale_factor <- max_weight / max_length

gg_combined_plot <- gg_combined_plot %>%
  mutate(Net_Yield_scaled = ifelse(Type == "Length", Net_Yield * scale_factor, Net_Yield),
         SE_scaled = ifelse(Type == "Length", SE * scale_factor, SE))

ggplot(gg_combined_plot, aes(x = factor(Treatment), fill = Type)) +
  geom_col(aes(y = Net_Yield_scaled), 
           position = position_dodge(width = 0.4), 
           color = "black", width = 0.3) +
  geom_errorbar(aes(ymin = Net_Yield_scaled, ymax = Net_Yield_scaled + SE_scaled),
                width = 0.2, position = position_dodge(width = 0.4)) +
  scale_fill_grey(start = 1, end = 0) +
  scale_y_continuous(
    name = "Net yield (g)", 
    sec.axis = sec_axis(~ . / scale_factor, name = "Net yield (cm)")
  ) +
  labs(
    title = expression("Net yield of " * italic("Gracilaria gracilis") * " by Treatment"),
    x = "Initial total length",
    fill = ""
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.line = element_line(color = "black", size = 0.5),
    axis.ticks = element_line(color = "black"),
    legend.position = "top"
  )


```

New models without week and control filtered out

```{r}

# GG Average Length (Treatment significant)

gg_filtered_length <- gg_avg_length_per_week_by_treatment %>%
  filter(Treatment %in% c(5, 10))

# Shapiro-Wilk normality test to test normality of gg avg length per treatment per week
shapiro.test(gg_filtered_length$Average) # normality assumption met

# Levene's Test to check for homogeneity of treatments in gg avg length per treatment per week
leveneTest(gg_filtered_length$Average, 
gg_filtered_length$Treatment, center = mean ) # assumption of equal variances met

# Model of average length per treatment
mGAL_f <-  lm(Average ~ Treatment, data = gg_filtered_length)

summary(mGAL_f)
aov(mGAL_f)
anova(mGAL_f)

# Adjusted R-squared:  0.8959

## GG Average Weight (Treatment insignificant)

gg_filtered_weight <- gg_avg_weight_per_week_by_treatment %>%
  filter(Treatment %in% c(5, 10))

# Shapiro-Wilk normality test to test normality of gg avg weight per treatment per week
shapiro.test(gg_filtered_weight$Average) # assumption of normality met

# Levene's Test to check for homogeneity of treatments in gg avg weight per treatment per week
leveneTest(Average ~ Treatment, center = mean, data = gg_filtered_weight) # assumption of equal variances met

# Model of average weight per treatment 
mGAW_f <-  lm(Average ~ Treatment, data = gg_filtered_weight)

summary(mGAW_f)
anova(mGAW_f)

# Adjusted R-squared:  -0.1462 

# GG DGR Length - Pearson correlation

gg_filtered_rgr_length <- gg_rgr_length %>%
  filter(Treatment %in% c(5, 10))

# Pearson product-moment correlation coefficient

cor.test(gg_filtered_rgr_length$RGR_Length,as.numeric( gg_filtered_rgr_length$Treatment), 
         method = "pearson")

# GG DGR Length - Model

gg_filtered_rgr_length <- gg_avg_rgr_length %>%
  filter(Treatment %in% c(5, 10))

# Shapiro-Wilk normality test to test normality of gg RGR Length per treatment per week

shapiro.test((gg_filtered_rgr_length$RGR_Length))
# normality assumption met

# Levene's Test to check for homogeneity of treatments in gg avg rgr length per treatment per week

leveneTest(RGR_Length ~ Treatment, center = mean, data = gg_filtered_rgr_length) # assumption of equal variances met    

# Model of rgr length  per treatment per week 

mGRL_f  <- lm(RGR_Length ~ Treatment, data = gg_filtered_rgr_length) 

summary(mGRL_f)
anova(mGRL_f)
aov(mGRL_f) 

# GG DGR Weight - Pearson correlation

gg_filtered_rgr_weight <- gg_avg_rgr_weight %>%
  filter(Treatment %in% c(5, 10))

# Pearson product-moment correlation coefficient

cor.test(gg_filtered_rgr_weight$RGR_Weight,as.numeric( gg_filtered_rgr_weight$Treatment), 
         method = "pearson")


# GG DGR Weight - Model

gg_filtered_rgr_weight <- gg_avg_rgr_weight %>%
  filter(Treatment %in% c(5, 10))

# Shapiro-Wilk normality test to test normality of gg RGR Weight per treatment per week

shapiro.test((gg_filtered_rgr_weight$RGR_Weight))
# normality assumption met

# Levene's Test to check for homogeneity of treatments in gg avg rgr weight per treatment per week

leveneTest(RGR_Weight ~ Treatment, center = mean, data = gg_filtered_rgr_weight) # assumption of equal variances met    

# Model of rgr weight  per treatment per week 

mGRW_f  <- lm(RGR_Weight ~ Treatment, data = gg_filtered_rgr_weight) 

summary(mGRW_f)
anova(mGRW_f)
aov(mGRW_f) 

# GL Average Length

gl_filtered_length <- gl_avg_length_per_week_by_treatment %>%
  filter(Treatment %in% c(5, 10))

# Shapiro-Wilk normality test 
shapiro.test(gl_filtered_length$Average) # normality assumption met

# Levene's Test to check for homogeneity of treatments 
leveneTest(Average ~ Treatment, center = mean, data = gl_filtered_length) # equal variances assumption met
       
# Model of average length per treatment

mLAL_f <-  lm(Average ~ Treatment, data = gl_filtered_length)

summary(mLAL_f)
anova(mLAL_f)

# GL Average Weight

gl_filtered_weight <- gl_avg_weight_per_week_by_treatment %>%
  filter(Treatment %in% c(5, 10))

# Shapiro-Wilk normality test 
shapiro.test(gl_filtered_weight$Average) # normality assumption met

# Levene's Test to check for homogeneity 
leveneTest(Average ~ Treatment, center = mean, data = gl_filtered_weight) # equal variances assumption met

# Model average weight per treatment per week
mLAW_f <-  lm(Average ~ Treatment, data = gl_filtered_weight)

summary(mLAW_f)
anova(mLAW_f)
aov(mLAW_f)

# GL DGR Length - Pearson correlation

gl_filtered_rgr_length <- gl_avg_rgr_length %>%
  filter(Treatment %in% c(5, 10))

# Pearson product-moment correlation coefficient

cor.test(gl_filtered_rgr_length$RGR_Length,as.numeric( gl_filtered_rgr_length$Treatment), 
         method = "pearson")

# GL DGR Weight - Pearson correlation

gl_filtered_rgr_weight <- gl_avg_rgr_weight %>%
  filter(Treatment %in% c(5, 10))

# Pearson product-moment correlation coefficient

cor.test(gl_filtered_rgr_weight$RGR_Weight,as.numeric(gl_filtered_rgr_weight$Treatment), 
         method = "pearson")

```







